---
title: "Practical Machine Learning Project"
author: "Anjali Sohoni"
date: "December 27, 2015"
output: html_document
---
## Introduction:
### Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

```{r}
library(htmltools)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(RColorBrewer)
library(rattle)
library(corrplot)

```
### Lets read the data which is already downloaded from given URL.

```{r}
trainRaw <- read.csv("pml-training.csv")
testRaw <- read.csv("pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
### Lets remove columns with NA values and columns which do not have corelation with given problem.

```{r}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 

classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]

```
### Create partitions in training set for for model development.
```{r}
set.seed(4444)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

### Train the model for random forest using crossvalidation.
```{r}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```

### Predict and compute confusion matrix. Also compute accuracy and out of sample error.
```{r}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)

accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```

### Now we predict for the given test data.
```{r}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result

corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")

treeModel <- rpart(classe ~ ., data=trainData, method="class")
fancyRpartPlot(treeModel)
```

### Create 1 file for each test case.

```{r}
answers <- result
pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("Results_all/problem_id_",i,".txt")
    write.table(x[i], file=filename, quote=FALSE,
                row.names=FALSE, col.names=FALSE)
  }
}
pml_write_files(answers)

```

